### 2022-04-24 LaTeX/report.tex
Updates in the report and in the proof

\documentclass[a4 paper, 12pt]{article}
\usepackage{amsmath, amsthm, amsfonts, amssymb, mathrsfs}
\usepackage{tikzit}
\usepackage{geometry}
\usepackage{hyperref}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{listings}
\usepackage[ruled,vlined, linesnumbered]{algorithm2e}

\input{style.tikzstyles}

% editorial comments in the text or in marginal notes
% 1st argument: initials of the person making the comment,
% 2nd argument: comment to insert
\long\def\ednote#1#2{\par\noindent\framebox{\begin{minipage}{0.99\linewidth}\linespread{.7}\footnotesize #1: #2\end{minipage}}\par}
\newcommand{\edmargin}[2]{\marginpar{\raggedright\linespread{.7}\tiny #1: #2}}

%__________________________________________
\newcommand{\ldb}{\mathopen{\ooalign{\makebox[.4em][l]{$\lbrack$}\cr\makebox[.4em][r]{$\lbrack$}\cr}}}                                                       
\newcommand{\rdb}{\mathclose{\ooalign{\makebox[.4em][l]{$\rbrack$}\cr\makebox[.4em][r]{$\rbrack$}\cr}}}

%__________________________________________

\hypersetup{
    colorlinks=true,
    linktoc=false,
    linkcolor=blue,
    citecolor=blue
}

%__________________________________________
% Colors definition
\definecolor{isa_red}{RGB}{255, 58, 71}
\definecolor{isa_blue}{RGB}{0, 103, 158}
\definecolor{isa_green}{RGB}{0, 157, 97}
\definecolor{isa_dark_green}{RGB}{0,131, 0}
\definecolor{isa_purple}{RGB}{174, 5, 238}
\definecolor{isa_dark_blue}{RGB}{26, 0, 253}

% Isabelle keywords
\newcommand{\apply}{{\color{isa_red}{apply}}}
\newcommand{\done}{{\color{isa_red}{done}}}
\newcommand{\datatype}{{\color{isa_blue}{datatype}}}
\newcommand{\inductive}{{\color{isa_blue}{inductive}}}
\newcommand{\abbreviation}{{\color{isa_blue}{abbreviation}}}
\newcommand{\thm}{{\color{isa_blue}{theorem}}}
\newcommand{\lm}{{\color{isa_blue}{lemma}}}
\newcommand{\fun}{{\color{isa_blue}{fun}}}
\renewcommand{\locale}{{\color{isa_blue}{locale}}}
\newcommand{\where}{{\color{isa_green}{where}}}
\renewcommand{\and}{{\color{isa_green}{and}}}
\newcommand{\fixes}{{\color{isa_green}{fixes}}}
\newcommand{\assumes}{{\color{isa_green}{assumes}}}
\newcommand{\shows}{{\color{isa_green}{shows}}}
\newcommand{\generic}[1]{{\color{isa_purple}{\textquotesingle#1}}}
\newcommand{\isa}[1]{\small\texttt{\\\noindent#1}}
\newcommand{\blue}[1]{{\color{isa_dark_blue}{#1}}}
\newcommand{\green}[1]{{\color{isa_dark_green}{#1}}}
\newcommand{\env}[1]{$(\!|$#1$|\!)$}

\theoremstyle{definition}
\newtheorem{isabelle}{Isabelle}

\lstdefinelanguage{isabelle}{%
    keywords=[1]{type_synonym,datatype,fun,abbreviation,definition,proof,lemma,theorem,corollary},
    keywordstyle=[1]\bfseries\color{isarblue},
    keywords=[2]{where,assumes,shows,and},
    keywordstyle=[2]\bfseries\color{isargreen},
    keywords=[3]{if,then,else,case,of,SOME,let,in,O},
    keywordstyle=[3]\color{isarblue},
}
\lstset{%
  language=isabelle,
  escapeinside={&}{&},
  columns=fixed,
  extendedchars,
  basewidth={0.5em,0.45em},
  basicstyle=\ttfamily,
  mathescape,
}

%__________________________________________

\def\NN{$\mathcal{N}~$}
\def\GG{$\mathcal{G}~$}
\def\VV{$\mathcal{V}~$}
\def\EE{$\mathcal{E}~$}

\renewcommand\qedsymbol{$\blacksquare$}

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem{definition}{Definition}
\newtheorem{corollary}{Corollary}[lemma]
\newtheorem{remark}{Remark}

%__________________________________________

% Node styles
\tikzstyle{white node}=[fill=none, draw=black, shape=circle, tikzit fill=white, tikzit draw=black]
\tikzstyle{red node}=[fill={rgb,255: red,232; green,118; blue,188}, draw=black, shape=circle, tikzit fill={rgb,255: red,232; green,118; blue,188}, tikzit draw=black]
\tikzstyle{blue node}=[fill={rgb,255: red,179; green,191; blue,255}, draw=black, shape=circle, tikzit fill={rgb,255: red,179; green,191; blue,255}, tikzit draw=black]
\tikzstyle{green node}=[fill={rgb,255: red,201; green,255; blue,189}, draw=black, shape=circle, tikzit fill={rgb,255: red,201; green,255; blue,189}, tikzit draw=black]

\tikzstyle{arrow}=[>]
\tikzstyle{new style 0}=[fill=none, draw=red, shape=circle, tikzit fill=none]
\tikzstyle{arrow}=[<--]
\tikzstyle{new edge style 2}=[- - -]

% Edge styles
\tikzstyle{right arrow}=[->]
\tikzstyle{right red arrow}=[->, draw=red, tikzit draw=red]
\tikzstyle{left arrow}=[<-]
\tikzstyle{left red arrow}=[<-, draw=red, tikzit draw=red]
\tikzstyle{right dashed arrow}=[->, dashed=1, dash pattern=on 1mm off 1mm, draw=red]
\tikzstyle{left dashed arrow}=[<-, dashed=1, dash pattern=on 1mm off 1mm, draw=red]
\tikzstyle{blue fill}=[-, fill={rgb,255: red,179; green,191; blue,255}, draw=none, tikzit draw=none, tikzit fill={rgb,255: red,213; green,207; blue,255}]
\tikzstyle{green fill}=[-, fill={rgb,255: red,201; green,255; blue,189}, draw=none, tikzit draw=none, tikzit fill={rgb,255: red,213; green,255; blue,208}]
\tikzstyle{red fill}=[-, fill={rgb,255: red,232; green,118; blue,188}, draw=none, tikzit draw=none, tikzit fill={rgb,255: red,232; green,118; blue,118}]

%-----------------------------------------------------------------

\begin{document}
\newgeometry{margin=2cm}
\begin{titlepage}
    \begin{figure}[!h]
        \centering
        \includegraphics[height = .1\textwidth]{/Users/vincent/Documents/MinesNancy/logoartem.png}
        \hspace{1cm}
        \includegraphics[height = .1\textwidth]{img/logoloria.jpg}
        \hspace{1cm}
        \includegraphics[height = .1\textwidth]{img/logoUL.png}
    \end{figure}
    \vspace{3cm}

    \begin{center}
%        \huge{Strongly connected components algorithms : parallelization and proofs}
%        \huge{Formal methods and assisted proofs: application to strongly connected components algorithms}
        \huge{Formal verification of an algorithm for computing strongly connected components}
    \end{center}
    \vspace{1cm}
    \begin{center}
        \large{Département Informatique --- Parcours Recherche \\ Tuteur : Stephan Merz}
    \end{center}
    \vspace{1 cm}
    \begin{center}
        Vincent TRÉLAT
    \end{center}
    \vspace{2 cm}
    \begin{center}
        \textit{\today}
    \end{center}
    \vspace{2 cm}
    \begin{center}
        ***
    \end{center}

\end{titlepage}
\restoregeometry
\pagebreak

\pagebreak

\tableofcontents
\pagebreak

\section{Introduction}
\subsection{Academic context}
This research work was carried out as part of my curriculum at the French \href{https://mines-nancy.univ-lorraine.fr}{École des Mines de Nancy}. All documents such as codes or source papers are available on a \href{https://github.com/VTrelat/Tarjan}{GitHub repository}.

\subsection{Formal methods}
Formal methods are a field of computer science related to mathematical logic and reasoning. The whole purpose of the discipline is to ensure by a logical proof that a given algorithm is not only correct on its domain of definition, but also to find -- or define -- that domain. Formal methods find applications in a variety of fields, both concrete, such as the railway industry or self-driving cars, and abstract, such as computational architecture.
\BlankLine
Although a formal proof lies first on paper, the real formalisation starts when proofs are mechanised in a proof assistant.

\ednote{sm}{%
  First, a stylistic remark: do not end lines in a \LaTeX\ document using \texttt{\textbackslash\textbackslash} (except in \texttt{tabular} environments or similar). Use a blank line to start a new paragraph: the style will define the appearance of a paragraph break.

  More importantly, formal methods should not be equated with theorem proving. They are really about giving precise, mathematical definitions to computer science concepts. Although the purpose of giving such definitions is to enable formal verification, many techniques besides theorem proving, such as model-based testing, run-time monitoring, model checking etc.\ are used.
}

\subsection{Isabelle (HOL)}
\begin{quote}
    Isabelle is a generic proof assistant. It allows mathematical formulas to be expressed in a formal language and provides tools for proving those formulas in a logical calculus.
\end{quote}
\begin{flushright}
    \href{https://isabelle.in.tum.de/}{isabelle.in.tum.de}
\end{flushright}
Isabelle is a really powerful proof assistant coming with a higher order logic (HOL) proving environment. Isabelle proofs are written in the Isar (``intelligible semi-automated reasoning'') language that is designed to make proofs readable and comprehensible for a mathematically inclined reader, with minimal overhead introduced by the formalism. In fact, ``assistant'' refers to the fact that the machine checks the proof provided by the user, in contrast to automatic theorem proving where the machine finds the proof itself. The tools for automation are intended to help the user write the proof at a conveniently high level, without needing to work at the level of a logical calculus, for example.

\subsection{Isabelle by example}
The following example is a good introduction to the use of Isabelle.\edmargin{sm}{I'd explain what 'v stands for.}

\pagebreak

\section{Formalisation}
\subsection{Strongly connected components}
\subsubsection{Directed graphs}\label{sec:directedgraphs}

\begin{definition}[Reachability]
    For two vertices $x$ and $y$ of \VV, the reachability relation is noted ``$\Rightarrow^*$'' such that $x \Rightarrow^* y$ iff $x$ can reach $y$ in \GG.
\end{definition}

\begin{remark}
    The relation $\Rightarrow^*$ is in fact the transitive closure of the binary relation $\Rightarrow$ defining edges in a graph.
\end{remark}

\begin{definition}[SCC]
    Let \GG = (\VV, \EE) be an directed graph.
$\mathcal{C} \subseteq \mathcal{V}$ is a strongly connected component (SCC) of \GG if:
\begin{equation*}
    \forall x, y \in \mathcal{C}, (x \Rightarrow^* y) \wedge (y \Rightarrow^* x)
\end{equation*}
\textit{i.e.} there is a path between every $x$ and $y$ in $\mathcal{C}$.
\BlankLine
$\mathcal{C}$ is maximal, or $\mathcal{C}$ is a maximal SCC of \GG if there is no other SCC containing $\mathcal{C}$, \textit{i.e.} if:
\begin{equation*}
    \forall \mathcal{X}, (\mathcal{C} \subseteq \mathcal{X}) \wedge (\forall x, y \in \mathcal{X}, (x \Rightarrow^* y) \wedge (y \Rightarrow^* x)) \Longrightarrow \mathcal{C} = \mathcal{X}
\end{equation*}
\end{definition}

\begin{definition}(Strong connectedness)
Let \GG = (\VV, \EE) be a directed graph. \GG is strongly connected if \VV is a SCC.
\end{definition}

\subsubsection{Examples}
\begin{figure}[!h]
    \centering
    \begin{subfigure}[t]{.49\textwidth}
        \ctikzfig{example2}
        \subcaption{Strongly connected graph}
    \end{subfigure}
    \begin{subfigure}[t]{.49\textwidth}
        \ctikzfig{example3}
        \subcaption{Not strongly connected graph}
    \end{subfigure}
    \caption{Basic example of what is a small SCC}
\end{figure}
\begin{figure}[!h]
    \ctikzfig{exampleSCC}
    \caption{Example of a graph where each colored set of node is a -- maximal -- SCC\label{fig:exampleSCC}}
\end{figure}

\begin{figure}[!h]
    \ctikzfig{exampleSCCreduced}
    \caption{Reduced visualization of the graph represented in figure \ref{fig:exampleSCC}}
\end{figure}

\pagebreak

\section{A sequential set-based algorithm}
\subsection{Formalisation}
\begin{definition}[SCC mapping]
    In the following algorithm, the SCCs are progressively tracked in a collection of disjoint\edmargin{sm}{We may have $\mathcal{S}(v) = \mathcal{S}(w)$ for $v \neq w$?!} sets through a map $\mathcal{S} : \mathcal{V} \longrightarrow \mathcal{P}(\mathcal{V})$, where $\mathcal{P}(\mathcal{V})$ is the powerset of \VV, s.t. the following invariant is maintained:
    \begin{equation}\label{invariant}
        \forall v, w \in \mathcal{V}, w\in \mathcal{S}(v) \Longleftrightarrow \mathcal{S}(v) = \mathcal{S}(w)
    \end{equation}
\end{definition}

\begin{remark}
    In particular, $\forall v \in \mathcal{V}, v \in \mathcal{S}(v)$.
\end{remark}

\begin{definition}[SCC union]
    Let \textsc{Unite} be the function taking as parameters a map $\mathcal{S}$ as defined previously and two vertices $u$ and $v$ of $\mathcal{V}$ such that $\textsc{Unite}(\mathcal{S}, u, v)$ merges the two mapped sets $\mathcal{S}(u)$ and $\mathcal{S}(v)$ and maintains the invariant (\ref{invariant}) by updating the function $\mathcal{S}$.
\end{definition}
Let us give an example:\\
Let $\mathcal{V} = \{u,v,w\}$ such that there is the following mapping: $\mathcal{S}(u) = \{u\}$ and $\mathcal{S}(v) = \mathcal{S}(w) = \{v,w\}$.\\
Then, $\textsc{Unite}(\mathcal{S}, u, v) = \mathcal{S}(u) = \mathcal{S}(v) = \mathcal{S}(w) = \{u,v,w\}$.

\begin{definition}[Successors set for a node]
    Let $\mathcal{G} = (\mathcal{V}, \mathcal{E})$ and $v \in \mathcal{V}$. The set of successors of $v$ in \GG is \textsc{Post}($v$) such that:
    \begin{equation*}
        \forall w\in \textsc{Post}(v), (v,w)\in \mathcal{E}
    \end{equation*}
\end{definition}

\subsection{The algorithm}
See \cite{articleSCC} for the original paper.\\
\begin{algorithm}[H]\label{alg:seqsetbased}
    \SetAlgoLined
    \KwData{A graph \GG = (\VV, \EE), a starting node $v_0$\;}
    \KwResult{A partition \texttt{SCCs} of \VV where each element of \texttt{SCCs} is a maximal set of strongly connected components of \GG\;}
    Initialize an empty set \texttt{EXPLORED}\;
    Initialize an empty set \texttt{VISITED}\;
    Initialize an empty stack \texttt{R}\;
    setBased($v_0$)\;
    \SetKwProg{Function}{function}{}{}
    \Function{setBased: $v \in \mathcal{V} \rightarrow \texttt{None}$}{
        $\texttt{VISITED} := \texttt{VISITED} \cup \{v\}$\;
        \texttt{R.push($v$)}\;
        \ForEach{$w\in \texttt{POST(v)}$}{
            \If{$w\in \texttt{EXPLORED}$}{
                continue\;
            }
            \ElseIf{$w \notin \texttt{VISITED}$}{
                setBased($w$)\;
            }
            \Else{
                \While{$\mathcal{S}(v) \neq \mathcal{S}(w)$}{
                    $r := \texttt{R.pop()}$\;
                    $\texttt{UNITE}(\mathcal{S}, r, \texttt{R.top()})$\;
                }
            }
        }
        \If{$v = \texttt{R.top()}$}{
            \textbf{report SCC} $\mathcal{S}(v)$\;
            $\texttt{EXPLORED} := \texttt{EXPLORED} \cup \mathcal{S}(v)$\;
            \texttt{R.pop()}\;
        }
    }

    \caption{Sequential set-based SCC algorithm}
\end{algorithm}
\ednote{sm}{%
  The algorithm should also explain how $\mathcal{S}$ is initialized.
}

\pagebreak

\subsection{Informal proof}
Note that this proof is said informal only because it is not checked by a mechanized proof assistant. Both logical and mathematical arguments developed below are absolutely relevant.

\begin{lemma}(First invariant)\label{lemma:disjointness}
    \begin{equation*}
        \forall x, y \in \texttt{R}, x\neq y \implies \mathcal{S}(x) \cap \mathcal{S}(y) = \varnothing
    \end{equation*}
    Note the misuse of the set notation $x, y \in \texttt{R}$ which just means that $x$ and $y$ are in the stack \texttt{R}.
\end{lemma}

\begin{proof}
    Let $x \in \mathcal{V}$ be the following node to be visited during the execution of the algorithm \ref{alg:seqsetbased}: $x$ is pushed in \texttt{R}. Let $y \in \textsc{Post}(x)$.\edmargin{sm}{The lemma assumes $y \in \texttt{R}$, is $y \in \textsc{Post}(x)$ an additional assumption? Why is it justified?} There are two cases:\\
    \begin{itemize}
        \item y has not been visited yet, \textit{i.e.} $y \notin \textsc{Visited}$. Thus, a DFS-like traversal is performed from $y$, so $y$ is pushed in \texttt{R} and $\mathcal{S}(y) = \{y\}$ because $y$ is alone in its equivalence class for the moment since it has not been visited yet.\\
        Therefore, $\mathcal{S}(x) \cap \mathcal{S}(y) = \varnothing$.\edmargin{sm}{I don't see an argument why $y \notin \mathcal{S}(x)$?}

        \item y has already been visited, \textit{i.e.} $y \in \textsc{Visited}$. Then, $y$ was already pushed in \texttt{R} before $x$. Let $(x_i)_{1\leq i \leq n}$ be the first nodes of the stack s.t. $x_0 = x$ and $x_n = y$.\edmargin{sm}{Should explain the order of the stack.}\\
        In order to avoid writing $\texttt{R} = [\ldots, y, \ldots, x]$, let us define $\widetilde{R}$ the stack containing the first $n$ nodes in \texttt{R}, s.t. $\widetilde{R} = [y, \ldots, x] = [x_n, \ldots, x_0]$.\\
        Let us consider the worst case, \textit{i.e.} when\edmargin{sm}{Why is this the worst case?}
        \begin{equation*}
            \forall~1 \leq i \leq n,~\mathcal{S}(x_i) = \{x_i\}
        \end{equation*}
        So, the while loop has to go down to $y$ because all partial SCCs are disjoint. As the length of the stack \texttt{R} is bounded by $|\mathcal{V}|$, the algorithm terminates.\\
        $x_0$ is first unstacked and both $\mathcal{S}(x_0)$ and $\mathcal{S}(\texttt{R.top()}) = \mathcal{S}(x_1)$ are then united. The current state of $\mathcal{S}$ and $\widetilde{R}$ is:\edmargin{sm}{Ambiguity between $\mathcal{S}$ as a mapping and as a set of equivalence classes.}
        \begin{equation*}
            \left\{
                \begin{array}{l}
                    \mathcal{S} = \{\{x_0, x_1\}, \{x_2\}, \ldots, \{x_n\}, \ldots\}\\
                    \widetilde{R} = [x_n, \ldots, x_1]\\
                \end{array}
            \right.
        \end{equation*}
        Then, $x_1$ is unstacked and $\mathcal{S}(x_1)$ and $\mathcal{S}(x_2)$ are then united, so that:
        \begin{equation*}
            \left\{
                    \begin{array}{l}
                        \mathcal{S} = \{\{x_0, x_1, x_2\}, \{x_3\}, \ldots, \{x_n\}, \ldots\}\\
                        \widetilde{R} = [x_n, \ldots, x_2]\\
                    \end{array}
                \right.
        \end{equation*}
        Finally (by induction), $\mathcal{S} = \{x_0, \ldots, x_n\}$ and $\widetilde{R} = [y]$, \textit{i.e.} $\mathcal{S}(x)=\mathcal{S}(y)$. It is important to notice that $x=x_0, x_1, \ldots, x_{n-1}$ are no longer in the stack, so this operation kept the invariant true.
    \end{itemize}
\end{proof}

\begin{lemma}
    \begin{equation*}
        \biguplus_{r \in \texttt{R}} \mathcal{S}(r) = \textsc{Live} := \textsc{Visited} \setminus \textsc{Explored}
    \end{equation*}
\end{lemma}

\begin{proof}
    The disjointness of all on-stack partial SCCs is given by lemma \ref{lemma:disjointness}. Nodes from $\textsc{Visited} \setminus \textsc{Explored}$ are in \texttt{R} because they are being processed. So, $\textsc{Live} \subseteq \texttt{R}$.\\
    By L.6-7 of algorithm \ref{alg:seqsetbased}, $\textsc{Visited} \subseteq \texttt{R}$.\\
    L.9-10 ensure that no explored node is pushed in \texttt{R}.\\
    L.24-25 keep the invariant by unstacking explored nodes from \texttt{R}, so $\texttt{R} \cap \textsc{Explored} = \varnothing$. Thus, $\texttt{R} = \textsc{Visited} \setminus \textsc{Explored} = \textsc{Live}$.
\end{proof}

\begin{corollary}[Strong version]\label{cor:cor1}
    \begin{equation*}
        \forall v \in \textsc{Live}, \exists!~r \in \texttt{R} \cap \mathcal{S}(v), \mathcal{S}(v) = \mathcal{S}(r)
    \end{equation*}
\end{corollary}
\begin{proof}
    Let $v \in \textsc{Live} = \displaystyle{\biguplus_{r \in \texttt{R}} \mathcal{S}(r)}$. $v$ is in a unique partial SCC $\mathscr{S} := \mathcal{S}(v)$. Because of lemma \ref{lemma:disjointness}, there cannot exist $x \neq y \in \texttt{R}$ s.t. $\mathcal{S}(x) = \mathcal{S}(y) = \mathscr{S}$. Thus, there exists a unique $x \in \texttt{R}$ s.t. $\mathcal{S}(x) = \mathscr{S}$ (and $x\in \texttt{R} \cap \mathscr{S}$).
\end{proof}

\begin{corollary}[Weak version]\label{cor:cor2}
    \begin{equation*}
        \forall v \in \mathcal{V}, \forall w \in \textsc{Post}(v), w \in \textsc{Live} \implies \exists w' \in \texttt{R}, \mathcal{S}(w') = \mathcal{S}(w)
    \end{equation*}
\end{corollary}

\begin{proof}
    Holds because of corollary \ref{cor:cor1}.
\end{proof}

\begin{remark}
    In the algorithm \ref{alg:seqsetbased}, this property is maintained by L.16-18. These lines also illustrate how the algorithm ``reads'' the SCCs. Corollary \ref{cor:cor2} shows that when the mapped representatives of the top two nodes of \texttt{R} are united (until $\mathcal{S}(w') = \mathcal{S}(v) = \mathcal{S}(w)$ since $w'$ has a path to $v$), then all united components are in the same SCC.
\end{remark}

\begin{remark}\label{rem:proof}
    Because \texttt{R} only contains exactly one representative for each partial SCC \edmargin{sm}{I think the fact that this representative is the lowest one in the stack also plays a role here?} (corollary \ref{cor:cor1}), after each step of the main loop -- \textit{i.e.} the DFS -- every partial SCC is actually maximal in the current set of visited nodes.
\end{remark}

\begin{theorem}
    The sequential algorithm \ref{alg:seqsetbased} is correct, \textit{i.e.} it returns a set of maximal SCCs.
\end{theorem}
\begin{proof}
    Holds by remark \ref{rem:proof}.
\end{proof}

\pagebreak

\subsection{Formal proof}
Since the informal proof seems to be convincing, the formal -- checked automatically -- proof can be written in Isabelle (HOL) based on the basis of the reasoning developed above.

\subsubsection{Environment setup}
The first definitions should be the different structures used in the algorithm. In particular, a record containing all the sets needed and described in the pseudo-code of algorithm \ref{alg:seqsetbased}. The environment has a generic type parameter, which is used to represent the type of the nodes in the graph (often integers):

\isa{
    {\color{isa_blue}record} \generic{v} env =\\
    $~~~\mathcal{S}$ :: "\generic{v} $\Rightarrow$ \generic{v} set"\\
    $~~~$explored :: "\generic{v} set"\\
    $~~~$visited :: "\generic{v} set"\\
    $~~~$sccs :: "\generic{v} set set"\\
    $~~~$stack :: "\generic{v} list"\\
}

The following lines define a graph structure and some useful natural relations:

\isa{
    {\locale} graph =\\
    $~~~$\fixes{} {\color{isa_dark_blue} vertices} :: "\generic{v} set" \and{} {\color{isa_dark_blue} successors} :: "\generic{v} $\Rightarrow$ \generic{v} set"\\
    $~~~$\assumes{} vfin: "finite {\color{isa_dark_blue}vertices}"\\
    $~~~$\and{} sclosed: "$\forall {\text{{\color{isa_dark_green}x}}} \in$ {\color{isa_dark_blue}vertices}. {\color{isa_dark_blue}successors} {\color{isa_dark_green}x} $\subseteq$ {\color{isa_dark_blue}vertices}"\\
}

The use of \texttt{successors} instead of an adjacency matrix, for instance, is a consequence of the fact that the algorithm is only concerned with the topological ordering of the nodes. For instance, nodes can represent integers, logical propositions or sets of states in a proving system for example.

\subsubsection{Reachability}
Now that graphs are defined, the reachability can be defined. Defining an edge is simply some rewriting of being a successor of one node.

\isa{
    \abbreviation{} edge \where\\
    $~~~$"{\color{isa_dark_blue}edge} {\color{isa_dark_green}x y} $\equiv$ {\color{isa_dark_green}y} $\in$ {\color{isa_dark_blue}successors} {\color{isa_dark_green}x}"
}
\BlankLine
\BlankLine
\BlankLine
\BlankLine

\noindent
Regarding the reachability binary relation, a choice has to be made since there are several ways to define it. In particular, there are two possible keywords, {\color{isa_blue}\texttt{inductive}} and {\color{isa_blue}\texttt{fun}}, respectively for an inductive or recursive definition. If both definitions are valid, the inductive one is kept for the following reasons. Although a recursive definition allows one to do some rewriting in the middle of terms, a recursive definition expresses both the positive and negative information\footnote{In this case, the positive information designates the fact of being reachable and the negative information designates the fact of not being reachable.} whereas the inductive one only expresses the positive information directly. Therefore, with an inductive definition, the negative information has to be proved. One would be right to argue that it would be more convenient to be able to tell without proving it that two nodes are not reachable from each other, but this does not interest us for the following. Another important point is that there is no datatype for a recursive definition, especially in this case with the transitive closure of the $\Rightarrow^*$ relation. Thus, the choice of the inductive definition is not a choice of simplicity but of necessity. Lastly, Isabelle will generate a simple inductive rule for the proofs split into the reflexive case, which stands in the definition, and the transitive case, which has to be proved.

\isa{
    \inductive{} reachable \where\\
    $~~~$reachable\_refl[iff]: "{\color{isa_dark_blue}reachable} {\color{isa_dark_green}x x}"\\
    $~{\text{|}}$ reachable\_succ[elim]: "$\ldb$edge {\color{isa_dark_green}x y};  {\color{isa_dark_blue}reachable} {\color{isa_dark_green}y z}$\rdb \Longrightarrow$ {\color{isa_dark_blue}reachable} {\color{isa_dark_green}x z}"\\
}

In order to be able to use those relations in the proofs later, it is essential to prove a list of lemmas, namely all the different natural properties that Isabelle cannot deduce\footnote{That is an abuse of language. The idea is for example that for the moment, there is no formal link between \texttt{edge} and \texttt{reachable}. The goal is to formalize it so Isabelle is logically able to both use and simplify some results in the proofs.} from nothing\footnote{There is actually a theorem fetcher that is particularly useful to find a basic set of lemmas.}. For instance, the following lemmas are essential.

\isa{
\lm{} succ\_reachable:\\
$~~~~~$\assumes{} "reachable {\color{isa_dark_blue}x y}" \and{} "edge {\color{isa_dark_blue}y z}"\\
$~~~~~$\shows "reachable {\color{isa_dark_blue}x z}"\\
$~~~~~${\color{isa_blue}using} assms {\color{isa_blue}by} induct auto\\
}
Mathematical writing: $\forall x, \forall y, \forall z, (x \Rightarrow^* y \wedge y \Rightarrow z) \Longrightarrow x \Rightarrow^* z$

\isa{
\lm{} reachable\_trans:\\
$~~~~~$\assumes{} y: "reachable \blue{x y}" \and{} z: "reachable \blue{y z}"\\
$~~~~~$\shows{} "reachable \blue{x z}"\\
$~~~~~${\color{isa_blue}using} assms {\color{isa_blue}by} induct auto\\
}
Mathematical writing: $\forall x, \forall y, \forall z, (x \Rightarrow^* y \wedge y \Rightarrow^* z) \Longrightarrow x \Rightarrow^* z$\\

As the formal proofs will enventually deal with strongly connected components, it is also essential to formally define SCCs. For the purpose of the proof, the property of being a SCC is called \texttt{sub\_scc} and being a \textit{maximal} SCC is called \texttt{is\_scc} :

\isa{
{\color{isa_blue}definition} is\_subscc \where\\
$~~~~~$"is\_subscc S $\equiv \forall$ x $\in$ S. $\forall$ y $\in$ S. reachable x y"\\
}
Mathematical writing: A set $S$ is a SCC if $\forall x \in S, \forall y \in S, x \Rightarrow^* y$

\isa{
{\color{isa_blue}definition} is\_scc \where\\
$~~~~~$"is\_scc S $\equiv$ S $\neq$ \{\} $\wedge$ is\_subscc S\\
$~~~~~\wedge$ ($\forall$ S'. S $\subseteq$ S' $\wedge$ is\_subscc S' $\longrightarrow$ S' = S)"\\
}
Mathematical writing: A non-empty SCC $S$ is maximal if for all SCC $S'$, $S \subseteq S' \Longrightarrow S'=S$\\

Once again, there are some lemmas to prove, such as telling Isabelle when an element can be added to a SCC, or that two vertices that are reachable from each other are in the same SCC, or that two SCCs having a common element are identical.\\

\subsubsection{Equivalence realtion and graph partition}

\isa{
{\color{isa_blue}function} unite :: "\generic{v} $\Rightarrow$ \generic{v} $\Rightarrow$ \generic{v} env $\Rightarrow$ \generic{v} env" \where{}\\
$~~~~~$"{\color{isa_dark_blue}unite} {\color{isa_dark_green}v w e} =\\
$~~~~~~~~~~$({\color{isa_blue}if} ($\mathcal{S}$ {\color{isa_dark_green}e v} = $\mathcal{S}$ {\color{isa_dark_green}e w}) {\color{isa_blue}then} {\color{isa_dark_green}e}\\
$~~~~~~~~~~${\color{isa_blue}else let} {\color{isa_dark_green}r} = hd(stack {\color{isa_dark_green}e});\\
$~~~~~~~~~~~~~~~~~~~${\color{isa_dark_green}r'}= hd(tl(stack {\color{isa_dark_green}e}));\\
$~~~~~~~~~~~~~~~~~~~${\color{isa_dark_green}joined} = $\mathcal{S}$ {\color{isa_dark_green}e r} $\cup~\mathcal{S}$ {\color{isa_dark_green}e r'};\\
$~~~~~~~~~~~~~~~~~~~${\color{isa_dark_green}e'}= {\color{isa_dark_green}e'} \env{stack := tl(stack {\color{isa_dark_green}e}), $\mathcal{S}$ := ($\lambda$n. {\color{isa_blue}if} n $\in$ {\color{isa_dark_green}joined} {\color{isa_blue}then} {\color{isa_dark_green}joined} {\color{isa_blue}else} $\mathcal{S}$ {\color{isa_dark_green}e n})}\\
$~~~~~~~~~~${\color{isa_blue}in} {\color{isa_dark_blue}unite} {\color{isa_dark_green}v w e'})"\\
$~~~~~${\color{isa_blue}by} pat\_completeness auto
}

\subsubsection{Ordering relation}\label{sec:orderingrelation}
In the proof, a precedence relation\footnote{In fact, a total order is being defined on stacks.} noted {\color{isa_dark_blue}$\bullet~\preceq~\bullet~\text{in}~\bullet$} will be needed on the stack. Let $x$ and $y$ be two nodes and $R$ be a stack. Informally, $x$ precedes $y$ in $R$ if $y$ was pushed in $R$ before $x$ (see \textsc{Figure} \ref{fig:stackorder}).

\begin{figure}[!h]
    \ctikzfig{stackorder}
    \caption{The ordering relation on stacks\label{fig:stackorder}}
\end{figure}

\begin{definition}[Ordering relation]
    Let $x$ and $y$ be two nodes and $xs$ be a stack.
    \begin{equation*}
        x \preceq y~\text{in}~xs \equiv \exists~h,~\exists~r,~(xs = h @ [x] @ r) \wedge (y \in [x]@r)
    \end{equation*}
\end{definition}

The idea is to later use the following property: if $x \preceq y~\text{in}~xs$, then $y \Rightarrow^* x$.\\
It is defined in Isabelle as follows:\\
\isa{
    {\color{isa_blue}definition} precedes ("\_ $\preceq$ \_ in \_" [100,100,100] 39) \where\\
$~~~~~$"{\color{isa_dark_green}x} $\preceq$ {\color{isa_dark_green}y} in {\color{isa_dark_green}xs} $\equiv$ $\exists${\color{isa_dark_green}h} {\color{isa_dark_green}r}. {\color{isa_dark_green}xs} = {\color{isa_dark_green}h} @ ({\color{isa_dark_green}x} \# {\color{isa_dark_green}r}) $\wedge$ {\color{isa_dark_green}y} $\in$ set ({\color{isa_dark_green}x} \# {\color{isa_dark_green}r})"\\
}

All the different properties (\textit{i.e.} lemmas) which follow this definition in the Isabelle implementation are detailed in the natural mathematical writing in \nameref{appendix} .\\

\subsubsection{Implementation of the algorithm}

Now that the environment is set up, the actual algorithm -- seen as a function -- can be implemented.\\
Since Isabelle does not support loops, the implementation will be split into two mutually recursive functions. The main function is called \texttt{dfs} and takes its name after the Depth First Search algorithm because the algorithm \ref{alg:seqsetbased} roughly consists in a deep traversal of a graph. The second function is called \texttt{dfss} and represents the \textit{while} loop of the algorithm \ref{alg:seqsetbased}. The two functions are mutually recursive because they recursively call each other. In particular, \texttt{dfss} will call either itself or \texttt{dfs}, depending on the case.

\isa{
    {\color{isa_blue}{function}} dfs :: "\generic{v} $\Rightarrow$ \generic{v} env $\Rightarrow$ \generic{v} env" \and{}\\
    $~~~~~~~~~~$dfss:: "\generic{v} $\Rightarrow$ \generic{v} set $\Rightarrow$ \generic{v} env $\Rightarrow$ \generic{v} env" \where\\
    "{\color{isa_dark_blue}dfs} {\color{isa_dark_green}v e} =\\
    $~~~~~$({\color{isa_blue}{let}} {\color{isa_dark_green}e1} = {\color{isa_dark_green}e}\env{visited := visited e $\cup$ \{{\color{isa_dark_green}v}\}, stack := ({\color{isa_dark_green}v} \# stack {\color{isa_dark_green}e})};\\
    $~~~~~~~~~~${\color{isa_dark_green}e'} = {\color{isa_dark_blue}dfss} {\color{isa_dark_green}v} (successors {\color{isa_dark_green}v}) {\color{isa_dark_green}e1}\\
    $~~~~~${\color{isa_blue}{in if}} {\color{isa_dark_green}v} = hd(stack {\color{isa_dark_green}e'})\\
    $~~~~~~~~~~${\color{isa_blue}{then}} {\color{isa_dark_green}e'}\env{sccs:=sccs {\color{isa_dark_green}e'} $\cup$ {$\mathcal{S}$ {\color{isa_dark_green}e' v}}, explored:=explored {\color{isa_dark_green}e'} $\cup$ ($\mathcal{S}$ {\color{isa_dark_green}e' v}), stack:=tl(stack {\color{isa_dark_green}e'})}\\
    $~~~~~~~~~~${\color{isa_blue}else} {\color{isa_dark_green}e'})"\\
    | "{\color{isa_dark_blue}dfss} {\color{isa_dark_green}v vs e} =\\
    $~~~~~$({\color{isa_blue}if} {\color{isa_dark_green}vs} = \{\} {\color{isa_blue}then} {\color{isa_dark_green}e}\\
    $~~~~~${\color{isa_blue}else} ({\color{isa_blue}let} {\color{isa_dark_green}w} = SOME {\color{isa_dark_green}x}. {\color{isa_dark_green}x} $\in$ {\color{isa_dark_green}vs}\\
    $~~~~~~~~~~${\color{isa_blue}in} ({\color{isa_blue}let} {\color{isa_dark_green}e'} = ({\color{isa_blue}if} {\color{isa_dark_green}w} $\in$ explored {\color{isa_dark_green}e} {\color{isa_blue}then} {\color{isa_dark_green}e}\\
    $~~~~~~~~~~~~~~~~~~~${\color{isa_blue}else if} {\color{isa_dark_green}w} $\notin$ visited {\color{isa_dark_green}e} {\color{isa_blue}then} {\color{isa_dark_blue}dfs} {\color{isa_dark_green}w e}\\
    $~~~~~~~~~~~~~~~~~~~${\color{isa_blue}else} unite {\color{isa_dark_green}v w e})\\
    $~~~~~~~~~~~~~~${\color{isa_blue}in} {\color{isa_dark_blue}dfss} {\color{isa_dark_green}v} ({\color{isa_dark_green}v} - \{{\color{isa_dark_green}w}\}) {\color{isa_dark_green}e'})))"\\
    $~~~${\color{isa_blue}by} pat\_completeness (force+)\\
}

The two last keywords require explanations as well : \texttt{pat\_completeness} stands for \textit{pattern completeness} and ensures that there is no missing patterns. The keyword \texttt{force} is used\footnote{\texttt{force} is more aggressive in instantiation and seems to find the right instance.} to help Isabelle know -- by proving it -- that both \texttt{dfs} and \texttt{dfss} are actually functions and that those functions are well defined with respect to the usual logical and mathematical meaning.

\pagebreak

\section{Appendix}\label{appendix}
\subsection{Some lemmas}
Those lemmas refer to the precedence relation introduced in \textsc{Section} \ref{sec:orderingrelation}.
\BlankLine
Let $x, y, z$ be three nodes, and let $xs, ys, zs$ be three lists of nodes representing stacks.\\
By abuse of language, if an element is on a stack, it is in the set of elements contained in the stack so the following statement can be written: $x$ is on $xs \Longleftrightarrow x \in xs$. However, $xs$ in not seen as the set representing $xs$ since an element may occur several times in a stack.\\
The operator $@$ denotes the concatenation and operates on two lists: $[x_0, \dots, x_n] @ [y_0, \dots, y_m] = [x_0, \dots, x_n, y_0, \dots, y_m]$.\\

\begin{flushleft}
    (i)
$x \preceq y~\text{in}~xs \Longrightarrow (x \in xs)\wedge(y \in xs)$
\end{flushleft}
\begin{flushleft}
    (ii)
$y \in [x] @ xs \Longrightarrow x \preceq y~\text{in}~([x] @ xs)$
\end{flushleft}

\begin{flushleft}
    (iii)
    $x \neq z \Longrightarrow (x \preceq y~\text{in}~([z] @ zs) \Longrightarrow x \preceq y ~\text{in}~zs)$
\end{flushleft}

\begin{flushleft}
    (iv)
    $(y \preceq x~\text{in}~([x] @ xs)) \wedge (x \notin xs) \Longrightarrow (x = y)$
\end{flushleft}

\begin{flushleft}
    (v)
    $y \in (ys @ [x]) \Longrightarrow y \preceq x~\text{in}~(ys @ [x] @ xs)$
\end{flushleft}

\begin{flushleft}
    (vi)
    $(x \preceq x~\text{in}~xs) = (x \in xs)$
\end{flushleft}

\begin{flushleft}
    (vii)
    $x \preceq y~\text{in}~xs \Longrightarrow x \preceq y~\text{in}~(ys @ xs)$
\end{flushleft}

\begin{flushleft}
    (viii)
    $x \notin ys \Longrightarrow (x \preceq y~\text{in}~(ys @ xs) \Longleftrightarrow x \preceq y~\text{in}~xs)$
\end{flushleft}

\begin{flushleft}
    (ix)
    $x \preceq y~\text{in}~xs \Longrightarrow x \preceq y~\text{in}~(xs @ ys)$
\end{flushleft}

\begin{flushleft}
    (x)
    $y \notin ys \Longrightarrow x \preceq y~\text{in}~(xs @ ys) \Longleftrightarrow x \preceq y~\text{in}~xs$
\end{flushleft}

\begin{flushleft}
    (xi)(transitivity)
    $(x \preceq y~\text{in}~xs) \wedge (y \preceq z~\text{in}~xs) \wedge \underset{\text{all elements of $xs$ are distinct}}{\underbrace{(\forall~0 \leq i < j \leq \text{length}(xs), xs[i] \neq xs[j])}} \Longrightarrow x \preceq z~\text{in}~xs$
\end{flushleft}

\begin{flushleft}
    (xi)(antisymmetry)
    $(x \preceq y~\text{in}~xs) \wedge (y \preceq x~\text{in}~xs) \wedge \underset{\text{all elements of $xs$ are distinct}}{\underbrace{(\forall~0 \leq i < j \leq \text{length}(xs), xs[i] \neq xs[j])}} \Longrightarrow x = y$
\end{flushleft}

% continue with precedes_append_right_iff

\pagebreak

\begin{thebibliography}{99}

\bibitem{tarjanMerz} R. Chen, C. Cohen, J.-J. Lévy, S. Merz, L. Théry, \emph{Formal Proofs of Tarjan’s Strongly Connected Components Algorithm in Why3, Coq and Isabelle}, 2019

\bibitem{thesisBloemen} V. Bloemen, A. Laarman, J. van de Pol, \emph{Multi-Core On-The-Fly SCC Decomposition}, 2016

\bibitem{articleSCC} V. Bloemen, \emph{Strong Connectivity and Shortest Paths for Checking Models}, 2019

\end{thebibliography}

\end{document}