### 2022-08-16 Isabelle/document/root.tex
refactored proofs and explanations

\documentclass[11pt,a4paper]{article}
\usepackage[T1]{fontenc}
\usepackage{isabelle,isabellesym}

% further packages required for unusual symbols (see also
% isabellesym.sty), use only when needed

%\usepackage{amssymb}
  %for \<leadsto>, \<box>, \<diamond>, \<sqsupset>, \<mho>, \<Join>,
  %\<lhd>, \<lesssim>, \<greatersim>, \<lessapprox>, \<greaterapprox>,
  %\<triangleq>, \<yen>, \<lozenge>

%\usepackage{eurosym}
  %for \<euro>

%\usepackage[only,bigsqcap,bigparallel,fatsemi,interleave,sslash]{stmaryrd}
  %for \<Sqinter>, \<Parallel>, \<Zsemi>, \<Parallel>, \<sslash>

%\usepackage{eufrak}
  %for \<AA> ... \<ZZ>, \<aa> ... \<zz> (also included in amssymb)

%\usepackage{textcomp}
  %for \<onequarter>, \<onehalf>, \<threequarters>, \<degree>, \<cent>,
  %\<currency>

% this should be the last package used
\usepackage{pdfsetup}

% urls in roman style, theory text in math-similar italics
\urlstyle{rm}
\isabellestyle{it}

% for uniform font size
%\renewcommand{\isastyle}{\isastyleminor}

\begin{document}

\title{A Set-based Algorithm for Computing Strongly Connected Components of a Graph}
\author{Stephan Merz and Vincent Trélat}
\maketitle

\begin{abstract}
  We prove the correctness of a sequential algorithm for computing strongly
  connected components (SCCs) of a graph due to Vincent Bloemen.
\end{abstract}

\tableofcontents

% sane default for proof documents
\parindent 0pt\parskip 0.5ex

% generated text of all theories
\input{session}

% optional bibliography ... doesn't appear to work
%\bibliographystyle{abbrv}
%\bibliography{root}

\begin{thebibliography}{9}
\bibitem{bloemen:strong}
Vincent Bloemen.
\newblock Strong Connectivity and Shortest Paths for Checking Models.
\newblock University of Twente, Enschede, The Netherlands, 2019.
  \url{https://ris.utwente.nl/ws/portalfiles/portal/122499728/thesis.pdf}

\bibitem{dijkstra:finding}
Edsger W. Dijkstra.
\newblock Finding the Maximum Strong Components in a Directed Graph.
\newblock In: Selected Writings in Computing -- A Personal Perspective,
  pp.\ 22-30.
  Texts and Monographs in Computer Science. Springer, 1982.

\bibitem{munro:efficient}
J. Ian Munro.
\newblock Efficient Determination of the Transitive Closure of a Directed Graph.
\newblock \emph{Information Processing Letters} 1(2):56-58, 1971.

\bibitem{tarjan:depth-first}
Robert Tarjan.
\newblock Depth first search and linear graph algorithms.
\newblock \emph{SIAM Journal on Computing}, 1972.
\end{thebibliography}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End: